% pyformex manual --- simple
% $Id$
% (C) B.Verhegghe


\section{\module{simple} --- Simple geometries}
\label{sec:simple}

\declaremodule{""}{simple}
\modulesynopsis{Simple geometries.}
\moduleauthor{'pyFormex project'}{'http://pyformex.berlios.de'}


This module contains some functions, data and classes for generating Formices with simple geometries. You need to \Code{import} this module in your scripts to have access to its contents.


\begin{funcdesc}{regularGrid}{x0,x1,nx}
  Create a regular grid between points \var{x0} and \var{x1}.

  x0 and x1 are n-dimensional points (usually 1D, 2D or 3D).
  The space between x0 and x1 is divided in nx equal parts. nx should have
  the same dimension as x0 and x1.
  The result is a rectangular grid of coordinates in an array with
  shape ( nx[0]+1, nx[1]+1, ..., n ).
\end{funcdesc}


\begin{funcdesc}{shape}{name}
  Return a Formex with one of the predefined named shapes.

  This is a convenience function returning a plex-2 Formex constructed
  from one of the patterns defined in the simple.Pattern dictionary.

  Currently, the following patterns are defined: line, angle, square, plus,
  cross, diamond, rtriangle, cube, star, star3d.
\end{funcdesc}
  

\begin{funcdesc}{line}{p1=[0.,0.,0.],p2=[1.,0.,0.],n=1}
  Return a Formex which is a line between two specified points.
    
  p1: first point, p2: second point
  The line is split up in n segments.
\end{funcdesc}


\begin{funcdesc}{circle}{a1=2.,a2=0.,a3=360.}
Creates a Formex which is a polygonal approximation to a circle or arc.

All points generated by this function lie on a circle with unit radius at the origin in the x-y-plane.

\var{a1} (the dash angle) is the angle enclosed by the begin and end point of each line segment.\\
\var{a2} (the module angle) is the angle enclosed between the start points of two subsequent line segments. 
If \Code{a2==0.0} is given, it will be taken equal to a1.\\
\var{a3} (the arc angle) is the total angle enclosed between the first point of the first segment and the en point of the last segment.

All angles are given in degrees and are measured in the direction from x to y-axis. The first point of the first segment is always on the x-axis.

The default values produce a full circle (approximately). If $\var{a3} < 360$, the result is an arc.
Large values of \var{a1} and \var{a2} result in polygones. Thus circle(120.) is an
equilateral triangle and circle(60.) is regular hexagone.

Remark that the default \Code{a2 == a1} produces a continuous line, while a2 > a1 results in a dashed line.

\end{funcdesc}


\begin{funcdesc}{polygon}{n}
Creates the circumference of a regular polygon with $n$ sides, inscribed in a circle with radius 1 and center at the origin. The first point lies on the axis 0. All points are in the (0,1) plane. 
The return value is a plex-2 Formex. 
This function is equivalent to \Code{circle(360./n)}.
\end{funcdesc}


\begin{funcdesc}{triangle}{}
Returns an equilateral triangle with side length 1. The first point is the origin, the second points is on the axis 0. The return value is a plex-3 Formex.
\end{funcdesc}


\begin{funcdesc}{sphere2}{nx,ny,r=1,bot=-90,top=90}
Return a sphere consisting of line elements.

A sphere with radius r is modeled by a regular grid of nx
longitude circles, ny latitude circles and their diagonals.
    
The 3 sets of lines can be distinguished by their property number:
1: diagonals, 2: meridionals, 3: horizontals.

The sphere caps can be cut off by specifying top and bottom latitude
angles (measured in degrees from 0 at north pole to 180 at south pole.
\end{funcdesc}

        
\begin{funcdesc}{sphere3}{nx,ny,r=1,bot=-90,top=90}
Return a sphere consisting of surface triangles

A sphere with radius r is modeled by the triangles fromed by a regular
grid of nx longitude circles, ny latitude circles and their diagonals.

The two sets of triangles can be distinguished by their property number:
1: horizontal at the bottom, 2: horizontal at the top.

The sphere caps can be cut off by specifying top and bottom latitude
angles (measured in degrees from 0 at north pole to 180 at south pole.
\end{funcdesc}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "pyformex"
%%% End: 
